---
title: "proj_section4"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
date: "2023-03-25"
---

Preprocess data.

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```


```{r}
set.seed(0)
library(MASS)
library(caret)
library(compiler)
library(Metrics)
enableJIT(level = 3)

df <- read.csv('unpopular_songs.csv')
df$explicit <- as.integer(factor(df$explicit))-1
split <- sample(1:length(df[,1]), length(df[,1])*0.8, replace=FALSE)
train <- df[split,]
test <- df[-split,]
```
KNNReg reproduction

```{r}
knn <- knnreg(train[,1:13], train$popularity, k = 50)
preds <- predict(knn, test[,1:13])
print(paste("repro knn corr =", cor(preds, test$popularity)))
print(paste("repro knn mse =", mean((preds- test$popularity)^2)))
print(paste("repro knn acc =", accuracy(test$popularity, preds)))
```

# PCA

```{r}
pca <- preProcess(train[,1:14], method=c('pca'))
pca
pca_train <- predict(pca, train[,])
pca_test <- predict(pca, test[,])
df_pca_train <- data.frame(pca_train[,5:14], train$popularity)
df_pca_test <- data.frame(pca_test[,5:14], test$popularity)
```

KNN Regression through the PCA

```{r}
fit <- knnreg(df_pca_train[,], train$popularity, k=50)
preds <- predict(fit, df_pca_test[,])
print(paste("mse =",mean((preds-test$popularity)^2)))
print(paste("correlation =",cor(preds,test$popularity)))
print(paste("acc =", accuracy(test$popularity, preds)))
```

This seems to somehow increase the correlation and decrease the mse, so PCA has increased the accuracy of the model and not decreased it.

# LDA

The dataframe is reduced due to the processing power required.

```{r}
small_df <- df[1:1000,1:14]
split <- sample(1:length(small_df[,1]), length(small_df[,1])*0.8, replace=FALSE)
small_train <- small_df[split,]
small_test <- small_df[-split,]
lda <- lda(small_train$popularity~., data=small_train)
lda_pred <- predict(lda, newdata=small_test, type='class')
print(paste("correct =",mean(lda_pred$class==small_test$popularity)))
print(paste("acc =", accuracy(small_test$popularity, lda_pred$class)))
```

This also seems to have increased the accuracy from phase 1 in particular. I don't know why each of these increased the accuracy of each worked.